---
title: "Thailand Precipitation and Dengue Analysis"
output: pdf_document
date: "2023-05-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


H0: The time series is non-stationary.
H1: The time series is stationary.

If the p-value from the test is less than alpha (significance level -> e.g. alpha = 0.05), then we can reject the null hypothesis and conclude that the time series is stationary.
If the p-value from the test is more than alpha, then we can´t reject the null hypothesis and conclude that the time series is non-stationary.


# Dickey-Fuller test: over one year


## Dickey-Fuller test 2006

***Incidence: 2006 Northern Thailand*** -> non-stationary (p-value = 0.8506)
```{r}
library(tseries)

sum2006_north <- data.frame(t(apply(Dengue_2006[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2006 <- c(sum2006_north)

adf.test(north_2006)
```

***Incidence: 2006 Northeastern Thailand*** -> non-stationary (p-value = 0.956)
```{r}
library(tseries)

sum2006_northeastern <- data.frame(t(apply(Dengue_2006[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2006 <- c(sum2006_northeastern)

adf.test(northeastern_2006)
```



***Incidence: 2006 Central Thailand*** -> non-stationary (p-value = 0.9014)
```{r}
library(tseries)

sum2006_central <- data.frame(t(apply(Dengue_2006[1:18,4:15], 2,sum))+t(apply(Dengue_2006[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2006 <- c(sum2006_central)

adf.test(central_2006)
```

***Incidence: 2006 Southern Thailand*** -> stationary (p-value = 0.01)
```{r}
library(tseries)

sum2006_south <- data.frame(t(apply(Dengue_2006[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2006 <- c(sum2006_south)

adf.test(south_2006)
```

***Precipitation: 2006 Southern Thailand*** -> non-stationary (p-value = 0.9333)
```{r}
library(tseries)

sum2006_south_pr <- data.frame(t(apply(Precipitation_2006[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2006_pr <- c(sum2006_south_pr)

adf.test(south_2006_pr)
```


## Dickey-Fuller test 2007

***Incidence: 2007 Northern Thailand*** -> non-stationary (p-value = 0.8823)
```{r}
library(tseries)

sum2007_north <- data.frame(t(apply(Dengue_2007[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2007 <- c(sum2007_north)

adf.test(north_2007)
```

***Incidence: 2007 Northeastern Thailand*** -> non-stationary (p-value = 0.9048)
```{r}
library(tseries)

sum2007_northeastern <- data.frame(t(apply(Dengue_2007[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2007 <- c(sum2007_northeastern)

adf.test(northeastern_2007)
```

***Incidence: 2007 Central Thailand*** -> non-stationary (p-value = 0.9527)
```{r}
library(tseries)

sum2007_central <- data.frame(t(apply(Dengue_2007[1:18,4:15], 2,sum))+t(apply(Dengue_2007[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2007 <- c(sum2007_central)

adf.test(central_2007)
```

***Incidence: 2007 Southern Thailand*** -> non-stationary (p-value = 0.8606)
```{r}
library(tseries)

sum2007_south <- data.frame(t(apply(Dengue_2007[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2007 <- c(sum2007_south)

adf.test(south_2007)
```

## Dickey-Fuller test 2008

***Incidence: 2008 Northern Thailand*** -> non-stationary (p-value = 0.9766)
```{r}
library(tseries)

sum2008_north <- data.frame(t(apply(Dengue_2008[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2008 <- c(sum2008_north)

adf.test(north_2008)
```

***Incidence: 2008 Northeastern Thailand*** -> non-stationary (p-value = 0.9197)
```{r}
library(tseries)

sum2008_northeastern <- data.frame(t(apply(Dengue_2008[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2008 <- c(sum2008_northeastern)

adf.test(northeastern_2008)
```

***Incidence: 2008 Central Thailand*** -> non-stationary (p-value = 0.9571)
```{r}
library(tseries)

sum2008_central <- data.frame(t(apply(Dengue_2008[1:18,4:15], 2,sum))+t(apply(Dengue_2008[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2008 <- c(sum2008_central)

adf.test(central_2008)
```

***Incidence: 2008 Southern Thailand*** -> non-stationary (p-value = 0.9054)
```{r}
library(tseries)

sum2008_south <- data.frame(t(apply(Dengue_2008[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2008 <- c(sum2008_south)

adf.test(south_2008)
```

## Dickey-Fuller test 2009

***Incidence: 2009 Northern Thailand*** -> non-stationary (p-value = 0.79
```{r}
library(tseries)

sum2009_north <- data.frame(t(apply(Dengue_2009[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2009 <- c(sum2009_north)

adf.test(north_2009)
```

***Incidence: 2009 Northeastern Thailand*** -> non-stationary (p-value = 0.8119)
```{r}
library(tseries)

sum2009_northeastern <- data.frame(t(apply(Dengue_2009[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2009 <- c(sum2009_northeastern)

adf.test(northeastern_2009)
```

***Incidence: 2009 Central Thailand*** -> non-stationary (p-value = 0.952)
```{r}
library(tseries)

sum2009_central <- data.frame(t(apply(Dengue_2009[1:18,4:15], 2,sum))+t(apply(Dengue_2009[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2009 <- c(sum2009_central)

adf.test(central_2009)
```

***Incidence: 2009 Southern Thailand*** -> non-stationary (p-value = 0.4625)
```{r}
library(tseries)

sum2009_south <- data.frame(t(apply(Dengue_2009[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2009 <- c(sum2009_south)

adf.test(south_2009)
```

## Dickey-Fuller test 2010

***Incidence: 20010 Northern Thailand*** -> non-stationary (p-value = 0.9501)
```{r}
library(tseries)

sum2010_north <- data.frame(t(apply(Dengue_2010[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2010 <- c(sum2010_north)

adf.test(north_2010)
```

***Incidence: 2010 Northeastern Thailand*** -> non-stationary (p-value = 0.8942)
```{r}
library(tseries)

sum2010_northeastern <- data.frame(t(apply(Dengue_2010[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2010 <- c(sum2010_northeastern)

adf.test(northeastern_2010)
```

***Incidence: 2010 Central Thailand*** -> non-stationary (p-value = 0.9153)
```{r}
library(tseries)

sum2010_central <- data.frame(t(apply(Dengue_2010[1:18,4:15], 2,sum))+t(apply(Dengue_2010[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2010 <- c(sum2010_central)

adf.test(central_2010)
```

***Incidence: 2010 Southern Thailand*** -> non-stationary (p-value = 0.9303)
```{r}
library(tseries)

sum2010_south <- data.frame(t(apply(Dengue_2010[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2010 <- c(sum2010_south)

adf.test(south_2010)
```

## Dickey-Fuller test 2011

***Incidence: 2011 Northern Thailand*** -> non-stationary (p-value = 0.8142)
```{r}
library(tseries)

sum2011_north <- data.frame(t(apply(Dengue_2011[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2011 <- c(sum2011_north)

adf.test(north_2011)
```

***Incidence: 2011 Northeastern Thailand*** -> non-stationary (p-value = 0.9374)
```{r}
library(tseries)

sum2011_northeastern <- data.frame(t(apply(Dengue_2011[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2011 <- c(sum2011_northeastern)

adf.test(northeastern_2011)
```

***Incidence: 2011 Central Thailand*** -> non-stationary (p-value = 0.99)
```{r}
library(tseries)

sum2011_central <- data.frame(t(apply(Dengue_2011[1:18,4:15], 2,sum))+t(apply(Dengue_2011[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2011 <- c(sum2011_central)

adf.test(central_2011)
```

***Incidence: 2011 Southern Thailand*** -> stationary (p-value = 0.01378)
```{r}
library(tseries)

sum2011_south <- data.frame(t(apply(Dengue_2011[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2011 <- c(sum2011_south)

adf.test(south_2011)
```

## Dickey-Fuller test 2012

***Incidence: 2012 Northern Thailand*** -> non-stationary (p-value = 0.9544)
```{r}
library(tseries)

sum2012_north <- data.frame(t(apply(Dengue_2012[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2012 <- c(sum2012_north)

adf.test(north_2012)
```

***Incidence: 2012 Northeastern Thailand*** -> non-stationary (p-value = 0.99)
```{r}
library(tseries)

sum2012_northeastern <- data.frame(t(apply(Dengue_2012[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2012 <- c(sum2012_northeastern)

adf.test(northeastern_2012)
```

***Incidence: 2012 Central Thailand*** -> non-stationary (p-value = 0.99)
```{r}
library(tseries)

sum2012_central <- data.frame(t(apply(Dengue_2012[1:18,4:15], 2,sum))+t(apply(Dengue_2012[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2012 <- c(sum2012_central)

adf.test(central_2012)
```

***Incidence: 2012 Southern Thailand*** -> non-stationary (p-value = 0.6082)
```{r}
library(tseries)

sum2012_south <- data.frame(t(apply(Dengue_2012[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2012 <- c(sum2012_south)

adf.test(south_2012)
```

## Dickey-Fuller test 2013

***Incidence: 2013 Northern Thailand*** -> non-stationary (p-value = 0.9626)
```{r}
library(tseries)

sum2013_north <- data.frame(t(apply(Dengue_2013[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2013 <- c(sum2013_north)

adf.test(north_2013)
```

***Incidence: 2013 Northeastern Thailand*** -> non-stationary (p-value = 0.9546)
```{r}
library(tseries)

sum2013_northeastern <- data.frame(t(apply(Dengue_2013[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2013 <- c(sum2013_northeastern)

adf.test(northeastern_2013)
```

***Incidence: 2013 Central Thailand*** -> non-stationary (p-value = 0.9847)
```{r}
library(tseries)

sum2013_central <- data.frame(t(apply(Dengue_2013[1:18,4:15], 2,sum))+t(apply(Dengue_2013[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2013 <- c(sum2013_central)

adf.test(central_2013)
```

***Incidence: 2013 Southern Thailand*** -> non-stationary (p-value = 0.6502)
```{r}
library(tseries)

sum2013_south <- data.frame(t(apply(Dengue_2013[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2013 <- c(sum2013_south)

adf.test(south_2013)
```

## Dickey-Fuller test 2014

***Incidence: 2014 Northern Thailand*** -> non-stationary (p-value = 0.9766)
```{r}
library(tseries)

sum2014_north <- data.frame(t(apply(Dengue_2008[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2014 <- c(sum2014_north)

adf.test(north_2014)
```

***Incidence: 2014 Northeastern Thailand*** -> non-stationary (p-value = 0.9539)
```{r}
library(tseries)

sum2014_northeastern <- data.frame(t(apply(Dengue_2014[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2014 <- c(sum2014_northeastern)

adf.test(northeastern_2014)
```

***Incidence: 2014 Central Thailand*** -> non-stationary (p-value = 0.6665)
```{r}
library(tseries)

sum2014_central <- data.frame(t(apply(Dengue_2014[1:18,4:15], 2,sum))+t(apply(Dengue_2014[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2014 <- c(sum2014_central)

adf.test(central_2014)
```

***Incidence: 2014 Southern Thailand*** -> non-stationary (p-value = 0.9256)
```{r}
library(tseries)

sum2014_south <- data.frame(t(apply(Dengue_2014[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2014 <- c(sum2014_south)

adf.test(south_2014)
```

## Dickey-Fuller test 2015

***Incidence: 2015 Northern Thailand*** -> non-stationary (p-value = 0.2958)
```{r}
library(tseries)

sum2015_north <- data.frame(t(apply(Dengue_2015[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2015 <- c(sum2015_north)

adf.test(north_2015)
```

***Incidence: 2015 Northeastern Thailand*** -> non-stationary (p-value = 0.9073)
```{r}
library(tseries)

sum2015_northeastern <- data.frame(t(apply(Dengue_2015[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2015 <- c(sum2015_northeastern)

adf.test(northeastern_2015)
```

***Incidence: 2015 Central Thailand*** -> non-stationary (p-value = 0.4513)
```{r}
library(tseries)

sum2015_central <- data.frame(t(apply(Dengue_2015[1:18,4:15], 2,sum))+t(apply(Dengue_2015[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2015 <- c(sum2015_central)

adf.test(central_2015)
```

***Incidence: 2015 Southern Thailand*** -> non-stationary (p-value = 0.9706)
```{r}
library(tseries)

sum2015_south <- data.frame(t(apply(Dengue_2015[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2015 <- c(sum2015_south)

adf.test(south_2015)
```

## Dickey-Fuller test 2016

***Incidence: 2016 Northern Thailand*** -> non-stationary (p-value = 0.9623)
```{r}
library(tseries)

sum2016_north <- data.frame(t(apply(Dengue_2016[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2016 <- c(sum2016_north)

adf.test(north_2016)
```

***Incidence: 2016 Northeastern Thailand*** -> non-stationary (p-value = 0.5795)
```{r}
library(tseries)

sum2016_northeastern <- data.frame(t(apply(Dengue_2016[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2016 <- c(sum2016_northeastern)

adf.test(northeastern_2016)
```

***Incidence: 2016 Central Thailand*** -> stationary (p-value = 0.01)
```{r}
library(tseries)

sum2016_central <- data.frame(t(apply(Dengue_2016[1:18,4:15], 2,sum))+t(apply(Dengue_2016[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2016 <- c(sum2016_central)

adf.test(central_2016)
```

***Incidence: 2016 Southern Thailand*** -> non-stationary (p-value = 0.2966)
```{r}
library(tseries)

sum2016_south <- data.frame(t(apply(Dengue_2016[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2016 <- c(sum2016_south)

adf.test(south_2016)
```

## Dickey-Fuller test 2017

***Incidence: 2017 Northern Thailand*** -> non-stationary (p-value = 0.99)
```{r}
library(tseries)

sum2017_north <- data.frame(t(apply(Dengue_2017[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2017 <- c(sum2017_north)

adf.test(north_2017)
```

***Incidence: 2017 Northeastern Thailand*** -> non-stationary (p-value = 0.824)
```{r}
library(tseries)

sum2017_northeastern <- data.frame(t(apply(Dengue_2017[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2017 <- c(sum2017_northeastern)

adf.test(northeastern_2017)
```

***Incidence: 2017 Central Thailand*** -> non-stationary (p-value = 0.99)
```{r}
library(tseries)

sum2017_central <- data.frame(t(apply(Dengue_2017[1:18,4:15], 2,sum))+t(apply(Dengue_2017[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2017 <- c(sum2017_central)

adf.test(central_2017)
```

***Incidence: 2017 Southern Thailand*** -> stationary (p-value = 0.01)
```{r}
library(tseries)

sum2017_south <- data.frame(t(apply(Dengue_2017[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2017 <- c(sum2017_south)

adf.test(south_2017)
```

## Dickey-Fuller test 2018

***Incidence: 2018 Northern Thailand*** -> non-stationary (p-value = 0.9702)
```{r}
library(tseries)

sum2018_north <- data.frame(t(apply(Dengue_2018[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2018 <- c(sum2018_north)

adf.test(north_2018)
```

***Incidence: 2018 Northeastern Thailand*** -> non-stationary (p-value = 0.9041)
```{r}
library(tseries)

sum2018_northeastern <- data.frame(t(apply(Dengue_2018[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2018 <- c(sum2018_northeastern)

adf.test(northeastern_2018)
```

***Incidence: 2018 Central Thailand*** -> non-stationary (p-value = 0.9668)
```{r}
library(tseries)

sum2018_central <- data.frame(t(apply(Dengue_2018[1:18,4:15], 2,sum))+t(apply(Dengue_2018[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2018 <- c(sum2018_central)

adf.test(central_2018)
```

***Incidence: 2018 Southern Thailand*** -> non-stationary (p-value = 0.9025)
```{r}
library(tseries)

sum2018_south <- data.frame(t(apply(Dengue_2018[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2018 <- c(sum2018_south)

adf.test(south_2018)
```

## Dickey-Fuller test 2019

***Incidence: 2019 Northern Thailand*** -> non-stationary (p-value = 0.9088)
```{r}
library(tseries)

sum2019_north <- data.frame(t(apply(Dengue_2019[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2019 <- c(sum2019_north)

adf.test(north_2019)
```

***Incidence: 2019 Northeastern Thailand*** -> non-stationary (p-value = 0.9593)
```{r}
library(tseries)

sum2019_northeastern <- data.frame(t(apply(Dengue_2019[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2019 <- c(sum2019_northeastern)

adf.test(northeastern_2019)
```

***Incidence: 2019 Central Thailand*** -> non-stationary (p-value = 0.9854)
```{r}
library(tseries)

sum2019_central <- data.frame(t(apply(Dengue_2019[1:18,4:15], 2,sum))+t(apply(Dengue_2019[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2019 <- c(sum2019_central)

adf.test(central_2019)
```

***Incidence: 2019 Southern Thailand*** -> non-stationary (p-value = 0.9873)
```{r}
library(tseries)

sum2019_south <- data.frame(t(apply(Dengue_2019[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2019 <- c(sum2019_south)

adf.test(south_2019)
```

## Dickey-Fuller test 2020

***Incidence: 2020 Northern Thailand*** -> non-stationary (p-value = 0.9088)
```{r}
library(tseries)

sum2020_north <- data.frame(t(apply(Dengue_2020[38:46,4:15], 2,sum))) # provinces_north = province_numbers[38:46,2]
north_2020 <- c(sum2020_north)

adf.test(north_2020)
```

***Incidence: 2020 Northeastern Thailand*** -> non-stationary (p-value = 0.9593)
```{r}
library(tseries)

sum2020_northeastern <- data.frame(t(apply(Dengue_2020[19:37,4:15], 2,sum))) # provinces_northeastern = province_numbers[19:37, 2]
northeastern_2020 <- c(sum2020_northeastern)

adf.test(northeastern_2020)
```

***Incidence: 2020 Central Thailand*** -> non-stationary (p-value = 0.9854)
```{r}
library(tseries)

sum2020_central <- data.frame(t(apply(Dengue_2020[1:18,4:15], 2,sum))+t(apply(Dengue_2020[47:62,4:15], 2,sum))) # provinces_central = province_numbers[c(1:18,47:62), 2]
central_2020 <- c(sum2020_central)

adf.test(central_2020)
```

***Incidence: 2020 Southern Thailand*** -> non-stationary (p-value = 0.9873)
```{r}
library(tseries)

sum2020_south <- data.frame(t(apply(Dengue_2020[63:76,4:15], 2,sum))) # provinces_south = province_numbers[63:76, 2]
south_2020 <- c(sum2020_south)

adf.test(south_2020)
```


# Dickey-Fuller test: 2006-2020

***Incidence: 2006-2020 Northern Thailand*** -> stationary (p-value = 0.01)
```{r}
library(tseries)

north <- c(sum2006_north, sum2007_north, sum2008_north, sum2009_north, sum2010_north, sum2011_north, sum2012_north, sum2013_north, sum2014_north, sum2015_north, sum2016_north, sum2017_north, sum2018_north, sum2019_north, sum2020_north)

adf.test(north)
```

***Incidence: 2006-2020 Northeastern Thailand*** -> stationary (p-value = 0.01)
```{r}
library(tseries)

northeastern <- c(sum2006_northeastern, sum2007_northeastern, sum2008_northeastern, sum2009_northeastern, sum2010_northeastern, sum2011_northeastern, sum2012_northeastern, sum2013_northeastern, sum2014_northeastern, sum2015_northeastern, sum2016_northeastern, sum2017_northeastern, sum2018_northeastern, sum2019_northeastern, sum2020_northeastern)

adf.test(northeastern)
```

***Incidence: 2006-2020 Central Thailand*** -> stationary (p-value = 0.01)
```{r}
library(tseries)

central <- c(sum2006_central, sum2007_central, sum2008_central, sum2009_central, sum2010_central, sum2011_central, sum2012_central, sum2013_central, sum2014_central, sum2015_central, sum2016_central, sum2017_central, sum2018_central, sum2019_central, sum2020_central)

adf.test(central)
```

***Incidence: 2006-2020 Southern Thailand*** -> stationary (p-value = 0.01)
```{r}
library(tseries)

south <- c(sum2006_south, sum2007_south, sum2008_south, sum2009_south, sum2010_south, sum2011_south, sum2012_south, sum2013_south, sum2014_south, sum2015_south, sum2016_south, sum2017_south, sum2018_south, sum2019_south, sum2020_south)

adf.test(south)
```


# Results Dickey-Fuller test 

## Results Dickey-Fuller test: incidences 

- dengue cases are in general non-stationary/seasonal over a year 
- exception: 2006 south, 2011 south, 2016 central, 2017 south
- the p-values are in general lower for Southern Thailand 
- dengue cases are stationary over the years (2006-2020)

## Results Dickey-Fuller test: precipitation

