---
title: "Test: GAM"
author: "Lena Koch"
date: "2023-06-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 

***load libraries***
```{r}
library(readxl)
library(nlme)
library(gam)
library(sdm)
library(raster)
library(ncdf4)
library(mgcv)
library(sf)
```

# 0 month lag

***attach dataframe***
```{r}
attach(GAM_0)
```

## Plot GAM 
```{r}
model0 <- gam(dengue~s(precipitation),family="quasipoisson")

plot.gam(model0)
summary(model0)
```


## GAM evaluation
```{r}
gam.check(model0)
```

in a “perfect” model: 
- the q-q plot should no curves
- the histogram have to be bell-shaped
- residuals values should be evenly distributed around 0
- the response vs fitted values should clusters around a one-one line
- p-value should not be too low -> indicates that k is too low, especially if edf is close to k
- gcv should be small 


## raster plot - precipitation per province
```{r}
linear_model <- lm(dengue~precipitation)
smooth_model <- gam(dengue~s(precipitation))

AIC(linear_model, smooth_model)


t2m_precipitation <- stack("pr_SEA22_MPI_rcp85_2021-2040_grid_daymean_monmean_swm_timmean_thai.nc")
names(t2m_precipitation) <- "precipitation"
t2m_precipitation

p0 <- raster::predict(t2m_precipitation, model0, type = "response")
```

```{r}
thailand <- raster::getData('GADM', country='THA', level=1) # in this case Thailand boundary is loaded

my.area <- extent(thailand) # the extent of a map describes the area visible on the map
my.p0 <- crop(p0, my.area) # crop returns a geographic subset of an object as specified by an Extent object 
my.p0.mask <- mask(my.p0, thailand) # mask function creates a MASK with values 1 and NULL, actually it plots 

# values only over Thailand, the rest are setting to zero

pdf(file="thailand_precipitation_dengue_0lag.pdf") # save the plot in pdf in your directory
plot(my.p0.mask, legend.width=1, legend.shrink=0.75, axes= FALSE, box=FALSE) 
title("Predicted values: Influence of precipitation on dengue (0 lag)", line = -1) 
 
plot(thailand, add=TRUE) 
dev.off()
```



# 1 month lag

***attach dataframe***
```{r}
attach(GAM_1)
```

## Plot GAM 
```{r}
model1 <- gam(dengue~s(precipitation),family="quasipoisson")

plot.gam(model1)
summary(model1)
```


## GAM evaluation
```{r}
gam.check(model1)
```

in a “perfect” model: 
- the q-q plot should no curves
- the histogram have to be bell-shaped
- residuals values should be evenly distributed around 0
- the response vs fitted values should clusters around a one-one line
- p-value should not be too low -> indicates that k is too low, especially if edf is close to k
- gcv should be small 


## raster plot - precipitation per province
```{r}
linear_model <- lm(dengue~precipitation)
smooth_model <- gam(dengue~s(precipitation))

AIC(linear_model, smooth_model)


t2m_precipitation <- stack("pr_SEA22_MPI_rcp85_2021-2040_grid_daymean_monmean_swm_timmean_thai.nc")
names(t2m_precipitation) <- "precipitation"
t2m_precipitation

p1 <- raster::predict(t2m_precipitation, model1, type = "response")
```

```{r}
thailand <- raster::getData('GADM', country='THA', level=1) # in this case Thailand boundary is loaded

my.area <- extent(thailand) # the extent of a map describes the area visible on the map
my.p1 <- crop(p1, my.area) # crop returns a geographic subset of an object as specified by an Extent object 
my.p1.mask <- mask(my.p1, thailand) # mask function creates a MASK with values 1 and NULL, actually it plots 

# values only over Thailand, the rest are setting to zero

pdf(file="thailand_precipitation_dengue_1lag.pdf") # save the plot in pdf in your directory
plot(my.p1.mask, legend.width=1, legend.shrink=0.75, axes= FALSE, box=FALSE) 
title("Predicted values: Influence of precipitation on dengue (1 lag)", line = -1) 
 
plot(thailand, add=TRUE) 
dev.off()
```



# 2 month lag

***attach dataframe***
```{r}
attach(GAM_2)
```

## Plot GAM 
```{r}
model2 <- gam(dengue~s(precipitation),family="quasipoisson")

plot.gam(model2)
summary(model2)
```


## GAM evaluation
```{r}
gam.check(model2)
```

in a “perfect” model: 
- the q-q plot should no curves
- the histogram have to be bell-shaped
- residuals values should be evenly distributed around 0
- the response vs fitted values should clusters around a one-one line
- p-value should not be too low -> indicates that k is too low, especially if edf is close to k
- gcv should be small 


## raster plot - precipitation per province
```{r}
linear_model <- lm(dengue~precipitation)
smooth_model <- gam(dengue~s(precipitation))

AIC(linear_model, smooth_model)


t2m_precipitation <- stack("pr_SEA22_MPI_rcp85_2021-2040_grid_daymean_monmean_swm_timmean_thai.nc")
names(t2m_precipitation) <- "precipitation"
t2m_precipitation

p2 <- raster::predict(t2m_precipitation, model2, type = "response")
```

```{r}
thailand <- raster::getData('GADM', country='THA', level=1) # in this case Thailand boundary is loaded

my.area <- extent(thailand) # the extent of a map describes the area visible on the map
my.p2 <- crop(p2, my.area) # crop returns a geographic subset of an object as specified by an Extent object 
my.p2.mask <- mask(my.p2, thailand) # mask function creates a MASK with values 1 and NULL, actually it plots 

# values only over Thailand, the rest are setting to zero

pdf(file="thailand_precipitation_dengue_2lag.pdf") # save the plot in pdf in your directory
plot(my.p2.mask, legend.width=1, legend.shrink=0.75, axes= FALSE, box=FALSE) 
title("Predicted values: Influence of precipitation on dengue (2 lag)", line = -1) 
 
plot(thailand, add=TRUE) 
dev.off()
```













