# Install beforehand

```{r}
install.packages("readxl")
library(readxl)
```

# Thailand Climate Data:

```{r}
daymin <- read_excel("Thailand.xlsx",sheet = "era5_t2m_daymin_annual_Thailand")
daymin
```

```{r}
daymax <- read_excel("Thailand.xlsx",sheet = "era5_t2m_daymax_annual_Thailand")
daymax
```

```{r}
tp <- read_excel("Thailand.xlsx",sheet = "era5_tp(mm)_annual_Thailand")
tp
```

# Dengue Data

Extracted from Gideon:

```{r}
Gideon <- read_excel("Dengue_Data.xlsx",sheet = "Gideon")
Gideon
```

# Questions for Week 1

## Question 1 (Lenny)

Identify which sub-districts are having the highest and the lowest temperature and total precipitation.

***Precipitation***

Step 1: Sort Data Set "Total Precipitation" in 2022 from highest to lowest

```{r}
tp_sorted = tp[order(tp$`2022`, decreasing = TRUE), ]
tp_sorted
```

Step Two: Give me the first six provinces with the highest precipitation and the lowest precipitation

```{r}
#sub districts with Highest Precipitation
tp_highest = head(tp_sorted)
tp_highest_list = tp_highest$ADM1_EN
tp_highest_list
```

```{r}
#sub districts with Lowest Precipitation
tp_lowest = tail(tp_sorted) 
tp_lowest_list = tp_lowest$ADM1_EN
tp_lowest_list
```

***Maximal Temperature***

Step 1: Sort Data Set "Max Temperature" in 2022 from highest to lowest

```{r}
daymax_sorted = daymax[order(daymax$`2022`, decreasing = TRUE), ]
daymax_sorted
```

Step Two: Give me the first six provinces with the highest max temperatur and the lowest max temperature

```{r}
#sub districts with Highest max Temperature
daymax_highest = head(daymax_sorted)
daymax_highest_list = daymax_highest$ADM1_EN
daymax_highest_list
```

```{r}
#sub districts with Lowest max temperature
daymax_lowest = tail(daymax_sorted) 
daymax_lowest_list = daymax_lowest$ADM1_EN
daymax_lowest_list
```

***Minimal Temperature***

Step 1: Sort Data Set "Min Temperature" in 2022 from highest to lowest

```{r}
daymin_sorted = daymin[order(daymin$`2022`, decreasing = TRUE), ]
daymin_sorted
```

Step Two: Give me the first six provinces with the highest min temperature and the lowest min temperature

```{r}
#sub districts with Highest min temperature
daymin_highest = head(daymin_sorted)
daymin_highest_list = daymin_highest$ADM1_EN
daymin_highest_list
daymin_highest

```

```{r}
#sub districts with Lowest min temperature
daymin_lowest = tail(daymin_sorted) 
daymin_lowest_list = daymin_lowest$ADM1_EN
daymin_lowest_list
```

***Highest Maximum and Minimal Temperature***

unique districts with the top highest min and max temperatures:

```{r}
temp_highest = unique(c(daymax_highest_list, daymin_highest_list))
temp_highest
```

***lowest Maximum and Minimal Temperature***

unique districts with the lowest min and max temperatures

```{r}
temp_lowest = unique(c(daymax_lowest_list, daymin_lowest_list))
temp_lowest
```

Summary: Sub-districts highest temperature

```{r}
temp_highest
```

sub-districts lowest temperature

```{r}
temp_lowest
```

sub-districts highest total precipitation.

```{r}
tp_highest_list
```

sub_districts lowest total precipitation

```{r}
tp_lowest_list
```

### Second Solution (Fiona)

***Province names***

```{r}
province_names = daymax[,2]
province_names
```

***Means over the years 2008-2022***

```{r}
#Mean over years 2008-2022 per province
tp_mean = apply(tp[,5:19], 1, mean)
tp_province_mean = cbind(province_names, tp_mean)
tp_province_mean

#Mean Min and Max Temperature per province
daymin_mean = apply(daymin[,5:19], 1, mean)
daymax_mean = apply(daymax[,5:19], 1, mean)

#Creation of a new dataframe with daymin and daymay per province
daymin_province_mean = cbind(daymin_mean, daymax_mean)
Temp_min_max = cbind(province_names, daymin_province_mean) 
Temp_min_max
```

***Mean Temperature of daymin and daymax per province***

```{r}
meantemp = apply(Temp_min_max[,2:3], 1, mean)
temp_mean =  cbind(province_names, meantemp)
temp_mean
```

***Provinces with highest mean temperature***

```{r}
highest_meantemp = temp_mean[head(order(temp_mean[,2], decreasing = TRUE)),]
highest_meantemp
highest_meantemp_provinces = highest_meantemp[,1]
highest_meantemp_provinces
```

***Provinces with lowest mean temperature***

```{r}
lowest_meantemp = temp_mean[tail(order(temp_mean[,2], decreasing = TRUE)),]
lowest_meantemp
lowest_meantemp_provinces = lowest_meantemp[,1]
lowest_meantemp_provinces
```

***mean total precipitation of all provinces***

```{r}
#mean of each province over all years
mean_tp = apply(tp[,5:19], 1, mean)
#add name of the provinces to the data frame
mean_tp = cbind(province_names, mean_tp)
colnames(mean_tp) = c("AMD1_EN", "meantp")
```

***provinces with highest and lowest total precipitation***

```{r}
#sort the provinces by mean total precipitation
mean_tp[order(mean_tp[,2], decreasing = TRUE),]

#Show 6 provinces each with highest and lowest total precipitation 
highest_meantp = head(mean_tp[order(mean_tp[,2], decreasing = TRUE),])
lowest_meantp  = head(mean_tp[order(mean_tp[,2], decreasing = FALSE),])

#show the provinces together with their mean total precipitation
highest_meantp
lowest_meantp
```

## Question 2 (Fiona)

Where are these districts located (east, west, north, south or central of the country)?

### Definition of the province locations in Thailand

```{r}
provinces_north = list("Chiang Mai", "Chiang Rai", "Lampang", "Lamphun", "Mae Hong Son", "Nan", "Phayao", "Phrae", "Uttaradit")

provinces_northeastern = list("Amnat Charoen", "Bueng Kan", "Buri Ram", "Chaiyaphum", "Kalasin", "Khon Kaen", "Loei", "Maha Sarakham", "Mukdahan", "Nakhon Phanom", "Nakhon Ratchasima", "Nong Bua Lam Phu", "Nong Khai", "Roi Et", "Sakon Nakhon", "Si Sa Ket", "Surin", "Ubon Ratchathani", "Udon Thani", "Yasothon")

provinces_south = list("Chumphon", "Krabi", "Nakhon Si Thammarat", "Narathiwat", "Pattani", "Phangnga", "Phatthalung", "Phuket", "Ranong", "Satun", "Songkhla", "Surat Thani", "Trang", "Yala")

provinces_central = list("Ang Thong", "Chachoengsao", "Chai Nat", "Chanthaburi", "Chon Buri", "Kamphaeng Phet", "Kanchanaburi", "Lop Buri", "Nakhon Nayok", "Nakhon Pathom", "Nakhon Sawan", "Nonthaburi", "Pathum Thani", "Phetchabun", "Phetchaburi", "Phichit", "Phitsanulok", "Phra Nakhon Si Ayutthaya", "Prachin Buri", "Prachuap Khiri Khan", "Ratchaburi", "Rayong", "Sa Keao", "Samut Prakan", "Samut Sakhon", "Samut Songkhram", "Saraburi", "Sing Buri", "Sukhothai", "Suphan Buri", "Tak", "Trat", "Uthai Thani")
```

###Location of the provinces with the highest mean temperature over the years 2008-2022

```{r}
#provinces in North Thailand
lowest_meantemp_provinces[which(lowest_meantemp_provinces %in% provinces_north)]

#provinces in Northeastern Thailand
lowest_meantemp_provinces[which(lowest_meantemp_provinces %in% provinces_northeastern)]

#provinces in Central Thailand
lowest_meantemp_provinces[which(lowest_meantemp_provinces%in% provinces_central)]

#provinces in South Thailand
lowest_meantemp_provinces[which(lowest_meantemp_provinces %in% provinces_south)]
```

### Location of the provinces with the lowest mean temperature over the years 2008-2022

```{r}
#provinces in North Thailand
highest_meantemp_provinces[which(highest_meantemp_provinces %in% provinces_north)]

#provinces in Northeastern Thailand
highest_meantemp_provinces[which(highest_meantemp_provinces %in% provinces_northeastern)]

#provinces in Central Thailand
highest_meantemp_provinces[which(highest_meantemp_provinces%in% provinces_central)]

#provinces in South Thailand
highest_meantemp_provinces[which(highest_meantemp_provinces %in% provinces_south)]
```

### Answers for Lennys Solution

####Location of the provinces with the highest temperature for 2022

```{r}
#provinces in North Thailand
temp_highest[which(temp_highest %in% provinces_north)]

#provinces in Northeastern Thailand
temp_highest[which(temp_highest %in% provinces_northeastern)]

#provinces in Central Thailand
temp_highest[which(temp_highest %in% provinces_central)]

#provinces in South Thailand
temp_highest[which(temp_highest %in% provinces_south)]
```

The provinces with the highest temperature are not located in north and northeastern Thailand. All provinces with the highest temperature except of Phuket are located in central Thailand. Phuket is located in the south.

#### Location of the provinces with the lowest temperature

```{r}
#provinces in North Thailand
temp_lowest[which(temp_lowest %in% provinces_north)]

#provinces in Northeastern Thailand
temp_lowest[which(temp_lowest %in% provinces_northeastern)]

#provinces in Central Thailand
temp_lowest[which(temp_lowest %in% provinces_central)]

#provinces in South Thailand
temp_lowest[which(temp_lowest %in% provinces_south)]
```

The provinces with the lowest temperature are mostly located in north and central Thailand.

#### Location of the provinces with the highest total precipitation

```{r}
#provinces in North Thailand
tp_highest_list[which(tp_highest_list %in% provinces_north)]

#provinces in Northeastern Thailand
tp_highest_list[which(tp_highest_list %in% provinces_northeastern)]

#provinces in Central Thailand
tp_highest_list[which(tp_highest_list %in% provinces_central)]

#provinces in South Thailand
tp_highest_list[which(tp_highest_list %in% provinces_south)]
```

The provinces with the highest total precipitation are mostly located in south Thailand. One of the provinces with the highest total precipitation is located in central Thailand.

#### Location of the provinces with the lowest total precipitation

```{r}
#provinces in North Thailand
tp_lowest_list[which(tp_lowest_list %in% provinces_north)]

#provinces in Northeastern Thailand
tp_lowest_list[which(tp_lowest_list %in% provinces_northeastern)]

#provinces in Central Thailand
tp_lowest_list[which(tp_lowest_list %in% provinces_central)]

#provinces in South Thailand
tp_lowest_list[which(tp_lowest_list %in% provinces_south)]
```

The provinces with the lowest precipitation are not located in north and south Thailand. 2/3 of them are located in central Thailand, 1/3 in northeastern Thailand.

## Question 3 (Lena)

Are the values of total precipitation realistic, please make some literature review and/or check with available historical weather data.

### Change in total precipitation over the years in the provinces

```{r}
plot(x = colnames(tp[,5:19]), y = tp[1,5:19], type = "b", pch = 19, col = "green", xlab = "year", ylab = "total precipitation", main = "total precipitation 2008-2022", ylim = c(900,3500)); lines(colnames(tp[,5:19]), tp[44,5:19], type = "b", pch = 19, col = "orange");  lines(colnames(tp[,5:19]), tp[37,5:19], type = "b", pch = 19, col = "blue"); lines(colnames(tp[,5:19]), tp[75,5:19], type = "b", pch = 19, col = "red"); 

legend(x = "topright",          
       legend = c("Pattani (south)", "Nan (north)", "Nakhon Phanom (northeastern)", "Bangkok (central)"),  
       lty = c(1),           
       col = c(col= "red",col="blue",col="orange", col="green"),           
       lwd = 2)     
```

```{r}
##total precipitation in different parts of thailands (directions based on "map_provinces")

#north
plot(x = colnames(tp[,5:19]), y = tp[44,5:19], type = "b", pch = 19, col="blue", ylim = c(900,3500), xlab = "year", ylab = "total precipitation", main = tp[44,2])
plot(x = colnames(tp[,5:19]), y = tp[46,5:19], type = "b", pch = 19, col="blue", ylim = c(900,3500), xlab = "year", ylab = "total precipitation", main = tp[46,2])

#northeastern
plot(x = colnames(tp[,5:19]), y = tp[22,5:19], type = "b", pch = 19, col="orange", ylim = c(900,3500), xlab = "year", ylab = "total percipitation", main = tp[22,2])
plot(x = colnames(tp[,5:19]), y = tp[23,5:19], type = "b", pch = 19, col="orange", ylim = c(900,3500), xlab = "year", ylab = "total percipitation", main = tp[23,2])

#central
plot(x = colnames(tp[,5:19]), y = tp[1,5:19], type = "b", pch = 19, col="green", ylim = c(900,3500), xlab = "year", ylab = "total precipitation", main = tp[1,2])
plot(x = colnames(tp[,5:19]), y = tp[2,5:19], type = "b", pch = 19, col="green", ylim = c(900,3500), xlab = "year", ylab = "total precipitation", main = tp[2,2])

#south
plot(x = colnames(tp[,5:19]), y = tp[75,5:19], type = "b", pch = 19, col="red", ylim = c(900,3500), xlab = "year", ylab = "total precipitation", main = tp[75,2])
plot(x = colnames(tp[,5:19]), y = tp[70,5:19], type = "b", pch = 19, col="red", ylim = c(900,3500), xlab = "year", ylab = "total precipitation", main = tp[70,2])

```

```{r}
hist(tp$"2008", ylim=c(0,20), breaks=15, main = colnames(tp[,5]), xlab = "total precipitation")
hist(tp$"2009", ylim=c(0,20), breaks=15, main = colnames(tp[,6]), xlab = "total precipitation")
hist(tp$"2010", ylim=c(0,20), breaks=15, main = colnames(tp[,7]), xlab = "total precipitation")
hist(tp$"2011", ylim=c(0,20), breaks=15, main = colnames(tp[,8]), xlab = "total precipitation")
hist(tp$"2012", ylim=c(0,20), breaks=15, main = colnames(tp[,9]), xlab = "total precipitation")
hist(tp$"2013", ylim=c(0,20), breaks=15, main = colnames(tp[,10]), xlab = "total precipitation")
hist(tp$"2014", ylim=c(0,20), breaks=15, main = colnames(tp[,11]), xlab = "total precipitation")
hist(tp$"2015", ylim=c(0,20), breaks=15, main = colnames(tp[,12]), xlab = "total precipitation")
hist(tp$"2016", ylim=c(0,20), breaks=15, main = colnames(tp[,13]), xlab = "total precipitation")
hist(tp$"2017", ylim=c(0,20), breaks=15, main = colnames(tp[,14]), xlab = "total precipitation")
hist(tp$"2018", ylim=c(0,20), breaks=15, main = colnames(tp[,15]), xlab = "total precipitation")
hist(tp$"2019", ylim=c(0,20), breaks=15, main = colnames(tp[,16]), xlab = "total precipitation")
hist(tp$"2020", ylim=c(0,20), breaks=15, main = colnames(tp[,17]), xlab = "total precipitation")
hist(tp$"2021", ylim=c(0,20), breaks=15, main = colnames(tp[,18]), xlab = "total precipitation")
hist(tp$"2022", ylim=c(0,20), breaks=15, main = colnames(tp[,19]), xlab = "total precipitation")
```

## Question 4 (Niklas)

Check the temporal variation of Dengue cases/deaths.

```{r}
#load Dengue cases and death
countries_cases = read_excel("Dengue_Data.xlsx", sheet = "Dengue_cases")
countries_death = read_excel("Dengue_Data.xlsx", sheet = "Denge_deaths")


#clean data just for Thailand
cases = countries_cases[,c(1,3)]
death = countries_death[,c(1,3)]


#plot to see cases and death numbers over time
plot(cases, type = "b", pch = 19, main = "Dengue cases Thailand"); plot(death, type = "b", pch = 19, main = "Dengue deaths Thailand") 
```

The number of cases peak every 2-3 years, as well as the death count. It should be observed next if there is a correlation between the peaks and the climate data. Between 2015 and 2019 was a rather calm period with less outbreaks than before, which could also be correlated to climate data.
