# Install beforehand

```{r}
install.packages("readxl")
library(readxl)
```

# Thailand Climate Data:

```{r}
install.packages("readxl")
library(readxl)

Thailand_daymin <- read_excel("Thailand.xlsx",sheet = "era5_t2m_daymin_annual_Thailand")
Thailand_daymin
```

```{r}
Thailand_daymax <- read_excel("Thailand.xlsx",sheet = "era5_t2m_daymax_annual_Thailand")
Thailand_daymax
```

```{r}
Thailand_tp <- read_excel("Thailand.xlsx",sheet = "era5_tp(mm)_annual_Thailand")
Thailand_tp
```

# Dengue Data

Extracted from Gideon:

```{r}
Gideon <- read_excel("Dengue_Data.xlsx",sheet = "Gideon")
Gideon
```

# Questions for Week 1

## Question 1 (Lenny)

Identify which sub-districts are having the highest and the lowest temperature and total precipitation.

```{r}
which()
```













## Question 2 (Fiona)

Where are these districts located (east, west, north, south or central of the country)?













## Question 3 (Lena)

Are the values of total precipitation realistic, please make some literature review and/or check with available historical weather data.

```{r}
tp = Thailand_tp[, c(3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19)]

plot(tp, type = "b", pch = 19)
```


```{r}
hist(Thailand_tp$"2008", breaks=15)
hist(Thailand_tp$"2009", breaks=15)
hist(Thailand_tp$"2010", breaks=15)
hist(Thailand_tp$"2011", breaks=15)
hist(Thailand_tp$"2012", breaks=15)
hist(Thailand_tp$"2013", breaks=15)
hist(Thailand_tp$"2014", breaks=15)
hist(Thailand_tp$"2015", breaks=15)
hist(Thailand_tp$"2016", breaks=15)
hist(Thailand_tp$"2017", breaks=15)
hist(Thailand_tp$"2018", breaks=15)
hist(Thailand_tp$"2019", breaks=15)
hist(Thailand_tp$"2020", breaks=15)
hist(Thailand_tp$"2021", breaks=15)
hist(Thailand_tp$"2022", breaks=15)
```

## Question 4 (Niklas)

Check the temporal variation of Dengue cases/deaths.

```{r}
#load Dengue cases and death
cases = read_excel("Dengue_Data.xlsx", sheet = "Dengue_cases")
death = read_excel("Dengue_Data.xlsx", sheet = "Denge_deaths")


#clean data just for Thailand
cases_Thai = cases[,c(1,3)]
death_Thai = death[,c(1,3)]


#plot to see cases and death numbers over time
plot(cases_Thai, type = "b", pch = 19); plot(death_Thai, type = "b", pch = 19)
```

The number of cases peak every 2-3 years, as well as the death count. It should be observed next if there is a correlation between the peaks and the climate data. Between 2015 and 2019 was a rather calm period with less outbreaks than before, which could also be correlated to climate data.
