# Install beforehand

```{r}
install.packages("readxl")
library(readxl)
```

# Thailand Climate Data:

```{r}
install.packages("readxl")
library(readxl)

Thailand_daymin <- read_excel("Thailand.xlsx",sheet = "era5_t2m_daymin_annual_Thailand")
Thailand_daymin
```

```{r}
Thailand_daymax <- read_excel("Thailand.xlsx",sheet = "era5_t2m_daymax_annual_Thailand")
Thailand_daymax
```

```{r}
Thailand_tp <- read_excel("Thailand.xlsx",sheet = "era5_tp(mm)_annual_Thailand")
Thailand_tp
```

# Dengue Data

Extracted from Gideon:

```{r}
Gideon <- read_excel("Dengue_Data.xlsx",sheet = "Gideon")
Gideon
```

# Questions for Week 1

## Question 1 (Lenny)

Identify which sub-districts are having the highest and the lowest temperature and total precipitation.

```{r}
which()
```










## Question 2 (Fiona)

Where are these districts located (east, west, north, south or central of the country)?










## Question 3 (Lena)

Are the values of total precipitation realistic, please make some literature review and/or check with available historical weather data.











## Question 4 (Niklas)

Check the temporal variation of Dengue cases/deaths.
```{r}
# load the data of the Dengue cases

```

