---
title: "Heatmaps"
author: "Fiona Steinacker"
date: "2023-07-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Heatmaps

### North

```{r}
library(pheatmap)
library(ggplot2)

H_Incidence_North = pheatmap(as.matrix(t(Incidence_North)), main = "Dengue Incidence North", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE) 
ggsave("Heatmap Incidence North.png", plot = H_Incidence_North)

H_Precipitation_North = pheatmap(as.matrix(t(Precipitation_North)), main = "Precipitation North", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE) 
ggsave("Heatmap Precipitation North.png", plot = H_Precipitation_North)
```

### Northeast

```{r}
H_Incidence_Northeast = pheatmap(as.matrix(t(Incidence_Northeast)), main = "Dengue Incidence Northeast", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE)
ggsave("Heatmap Incidence Northeast.png", plot = H_Incidence_Northeast)

H_Precipitation_Northeast = pheatmap(as.matrix(t(Precipitation_Northeast)), main = "Precipitation Northeast", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE)
ggsave("Heatmap Precipitation Northeast.png", plot = H_Precipitation_Northeast)
```

### Central

```{r}
H_Incidence_Central = pheatmap(as.matrix(t(Incidence_Central)), main = "Dengue Incidence Central", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE) 
ggsave("Heatmap Incidence Central.png", plot = H_Incidence_Central)

H_Precipitation_Central = pheatmap(as.matrix(t(Precipitation_Central)), main = "Precipitation Central", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE)
ggsave("Heatmap Precipitation Central.png", plot = H_Precipitation_Central)
```

### South

```{r}
H_Incidence_South = pheatmap(as.matrix(t(Incidence_South)), main = "Dengue Incidence South", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE) 
ggsave("Heatmap Incidence South.png", plot = H_Incidence_South)

H_Precipitation_South = pheatmap(as.matrix(t(Precipitation_South)), main = "Precipitation South", scale = "column", border_color = "white", cluster_rows = FALSE, cluster_cols = FALSE)
ggsave("Heatmap Precipitation South.png", plot = H_Precipitation_South)
```


# Line pLots

## incidence line plots for every year between 2006 and 2020  

```{r}
months_numbers <- 1:12
colors_regions <- c("purple2", "orangered2", "mediumseagreen", "maroon2")
names_regions <- c("North", "Northeast", "Central", "South")
years <- c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)

for (i in 1:15) {
  par(mar = c(5.1, 4.1, 4.1, 8.1), xpd = TRUE)
  plot(x = months_numbers, y = Incidence_North[i,], type = "b", pch = 20, lwd = 2, col = colors_regions[1], xlab = "month", ylab = "Incidences", main = paste("Incidences", years[i]), ylim = c(0, 170))
  for (j in 2:4) {
    lines(months_numbers, get(paste0("Incidence_", names_regions[j]))[i,], type = "b", pch = 20, lwd = 2, col = colors_regions[j])
  }
  
  legend(x = "topright",
         inset = c(-0.3,0),
         legend = names_regions,
         lty = 1,
         col = colors_regions,
         lwd = 2)
  
}
```

## incidence line plot: 2006-2020 

```{r}
#png(file = "plot_incidence.png", width = 1200, height = 550)

plot(x = 1:180, y = c(Incidence_North[1,],Incidence_North[2,],Incidence_North[3,],Incidence_North[4,],Incidence_North[5,],Incidence_North[6,],Incidence_North[7,],Incidence_North[8,],Incidence_North[9,],Incidence_North[10,],Incidence_North[11,],Incidence_North[12,], Incidence_North[13,],Incidence_North[14,],Incidence_North[15,]), type = "l", pch = 20, col = "purple2", xlab = "Month", ylab = "Incidence", main = "Incidences 2006-2020", ylim = c(0,170), xaxt = "n", xaxs="i", yaxs="i"); 

axis(1, at=seq(12,180, by=12));

lines(1:180, c(Incidence_Northeast[1,],Incidence_Northeast[2,], Incidence_Northeast[3,],Incidence_Northeast[4,], Incidence_Northeast[5,],Incidence_Northeast[6,], Incidence_Northeast[7,],Incidence_Northeast[8,], Incidence_Northeast[9,],Incidence_Northeast[10,], Incidence_Northeast[11,],Incidence_Northeast[12,], Incidence_Northeast[13,],Incidence_Northeast[14,], Incidence_Northeast[15,]), type = "l", pch = 20, col = "orangered2"); 

lines(1:180, c(Incidence_Central[1,],Incidence_Central[2,], Incidence_Central[3,],Incidence_Central[4,], Incidence_Central[5,],Incidence_Central[6,], Incidence_Central[7,],Incidence_Central[8,], Incidence_Central[9,],Incidence_Central[10,], Incidence_Central[11,],Incidence_Central[12,], Incidence_Central[13,],Incidence_Central[14,], Incidence_Central[15,]), type = "l", pch = 20, col = "mediumseagreen"); 

lines(1:180, c(Incidence_South[1,],Incidence_South[2,], Incidence_South[3,],Incidence_South[4,], Incidence_South[5,],Incidence_South[6,], Incidence_South[7,],Incidence_South[8,], Incidence_South[9,],Incidence_South[10,], Incidence_South[11,],Incidence_South[12,], Incidence_South[13,],Incidence_South[14,], Incidence_South[15,]), type = "l", pch = 20, col = "maroon2");

legend(x = "topright", 
       cex = 1.3,
       legend = c("north", "northeast", "central", "south"),  
       lty = c(1),           
       col = c(col= "purple2",col="orangered2",col="mediumseagreen", col="maroon2"),
       lwd = 2) 

#dev.off()
```


## precipitation line plots for every year between 2006 and 2020

```{r}
months_numbers <- 1:12
colors_regions <- c("purple2", "orangered2", "mediumseagreen", "maroon2")
names_regions <- c("North", "Northeast", "Central", "South")
years <- c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)

for (i in 1:15) {
  par(mar = c(5.1, 4.1, 4.1, 8.1), xpd = TRUE)
  plot(x = months_numbers, y = Precipitation_North[i,], type = "b", pch = 20, lwd = 2, col = colors_regions[1], xlab = "month", ylab = "Precipitation", main = paste("Precipitation", years[i]), ylim = c(0, 600))
  for (j in 2:4) {
    lines(months_numbers, get(paste0("Precipitation_", names_regions[j]))[i,], type = "b", pch = 20, lwd = 2, col = colors_regions[j])
  }
  
  legend(x = "topright",
         inset = c(-0.3,0),
         legend = names_regions,
         lty = 1,
         col = colors_regions,
         lwd = 2)
}
```

## precipitation line plot: 2006-2020

```{r}
#png(file = "plot_precipitation1.png", width = 1500, height = 700)

plot(x = 1:180, y = c(Precipitation_North[1,],Precipitation_North[2,],Precipitation_North[3,],Precipitation_North[4,],Precipitation_North[5,],Precipitation_North[6,],Precipitation_North[7,],Precipitation_North[8,],Precipitation_North[9,],Precipitation_North[10,],Precipitation_North[11,],Precipitation_North[12,], Precipitation_North[13,],Precipitation_North[14,],Precipitation_North[15,]), type = "l", pch = 20, col = "purple2", xlab = "Month", ylab = "Precipitation", main = "Precipitation 2006-2020", ylim = c(0,600), xaxt = "n", xaxs="i", yaxs="i"); 

axis(1, at=seq(12,180, by=12))

lines(1:180, c(Precipitation_Northeast[1,],Precipitation_Northeast[2,], Precipitation_Northeast[3,],Precipitation_Northeast[4,], Precipitation_Northeast[5,],Precipitation_Northeast[6,], Precipitation_Northeast[7,],Precipitation_Northeast[8,], Precipitation_Northeast[9,],Precipitation_Northeast[10,], Precipitation_Northeast[11,],Precipitation_Northeast[12,], Precipitation_Northeast[13,],Precipitation_Northeast[14,], Precipitation_Northeast[15,]), type = "l", pch = 20, col = "orangered2");  

lines(1:180, c(Precipitation_Central[1,],Incidence_Central[2,], Precipitation_Central[3,],Precipitation_Central[4,], Precipitation_Central[5,],Precipitation_Central[6,], Precipitation_Central[7,],Precipitation_Central[8,], Precipitation_Central[9,],Precipitation_Central[10,], Precipitation_Central[11,],Precipitation_Central[12,], Precipitation_Central[13,],Precipitation_Central[14,], Precipitation_Central[15,]), type = "l", pch = 20, col = "mediumseagreen"); 

lines(1:180, c(Precipitation_South[1,],Precipitation_South[2,], Precipitation_South[3,],Precipitation_South[4,], Precipitation_South[5,],Precipitation_South[6,], Precipitation_South[7,],Precipitation_South[8,], Precipitation_South[9,],Precipitation_South[10,], Precipitation_South[11,],Precipitation_South[12,], Precipitation_South[13,],Precipitation_South[14,], Precipitation_South[15,]), type = "l", pch = 20, col = "maroon2");

legend(x = "topright", 
       cex = 1.3,
       legend = c("North", "Northeast", "Central", "South"),  
       lty = c(1),           
       col = c(col= "purple2",col="orangered2",col="mediumseagreen", col="maroon2"),
       lwd = 2)   

#dev.off()
```


# Correlation of every province and year including Lag

```{r}
Final = c(1:30)

for (p in 1:76) {
  
  b = 5:16
  a = b-2
  n = 2006

  x = as.vector(Precipitation[p, b], mode = "double")
  y = as.vector(Incidence_combined[p, a], mode = "double")
  y1 = as.vector(Incidence_combined[p, a+1], mode = "double")
  y2 = as.vector(Incidence_combined[p, a+2], mode = "double")

  Cor = data.frame(c("Pearson","Spearman"),
                 c(cor(x,y, method = "pearson"), cor(x,y, method = "spearman")),
                 c(cor(x,y1, method = "pearson"), cor(x,y1, method = "spearman")),
                 c(cor(x,y2, method = "pearson"), cor(x,y2, method = "spearman")))

  Cor = cbind(n, Cor)

  colnames(Cor) = c(paste("year",Incidence_combined$Districts_names[p]), "Correlation method","no Lag", "1 month Lag", "2 month Lag")

  i = 1
  n = 2006

  repeat {
    b = 5:16 + 12*i
    a = b-2
    n = n + 1
    
    x = as.vector(Precipitation[p, b], mode = "double")
    y = as.vector(Incidence_combined[p, a], mode = "double")
    y1 = as.vector(Incidence_combined[p, a+1], mode = "double")
    y2 = as.vector(Incidence_combined[p, a+2], mode = "double")
    
    add = data.frame(c("Pearson","Spearman"),
                   c(cor(x,y, method = "pearson"), cor(x,y, method = "spearman")),
                   c(cor(x,y1, method = "pearson"), cor(x,y1, method = "spearman")),
                   c(cor(x,y2, method = "pearson"), cor(x,y2, method = "spearman")))
    add = cbind(n, add)
    
    colnames(add) = c(paste("year",Incidence_combined$Districts_names[p]),"Correlation method","no Lag", "1 month Lag", "2 month Lag")
    
    Cor = rbind(Cor, add)
    
    i = i + 1
    if (i == 14) {
      b = 5:16 + 12*i
      a = b-2
      n = n + 1
      
      x = as.vector(Precipitation[p, b], mode = "double")
      y = as.vector(Incidence_combined[p, a], mode = "double")
  
      x1 = x[-12]
      y1 = as.vector(Incidence_combined[p, a[-12] + 1], mode = "double")
  
      x2 = x[-(11:12)]
      y2 = as.vector(Incidence_combined[p, a[-(11:12)] + 2], mode = "double")
  
      add = data.frame(c("Pearson","Spearman"),
                   c(cor(x,y, method = "pearson"), cor(x,y, method = "spearman")),
                   c(cor(x1,y1, method = "pearson"), cor(x1,y1, method = "spearman")),
                   c(cor(x2,y2, method = "pearson"), cor(x2,y2, method = "spearman")))
  
      add = cbind(n, add)
  
      colnames(add) = c(paste("year",Incidence_combined$Districts_names[p]), "Correlation method","no Lag", "1 month Lag", "2 month Lag")
  
      Cor = rbind(Cor, add)
      
      break
    }
  }
  Final = cbind(Final, Cor)
  
  if (p == 1) {
    Final = Final[,-1]
  }
}
```

## Heatmap data for correlation of every year and province

```{r}
# choose every Pearson/Spearman correlation row of every year
Correlation_method = c("Pearson", "Spearman")

for (i in Correlation_method) {
  heatmap_data = Final[which(Final[,2] == i), ]
  rownames(heatmap_data) = heatmap_data[,1]
  
  # choosing lag method
  Lag = c("no_lag", "1_month_lag", "2_month_lag")
  
  ## vector for column selection
  b = c()
  for (n in 0:75){
    b = c(b, 3+5*n)
  }
  
  for (j in 1:3) {
    assign(paste0("heatmap_data_", Lag[j], "_", i), heatmap_data[,b-1+j])
  }
  
}

colnames(heatmap_data_no_lag_Pearson) = province_names[-which(province_names[,1] == "Bungkan") ,1]
colnames(heatmap_data_no_lag_Spearman) = province_names[-which(province_names[,1] == "Bungkan") ,1]
colnames(heatmap_data_1_month_lag_Pearson) = province_names[-which(province_names[,1] == "Bungkan") ,1]
colnames(heatmap_data_1_month_lag_Spearman) = province_names[-which(province_names[,1] == "Bungkan") ,1]
colnames(heatmap_data_2_month_lag_Pearson) = province_names[-which(province_names[,1] == "Bungkan") ,1]
colnames(heatmap_data_2_month_lag_Spearman) = province_names[-which(province_names[,1] == "Bungkan") ,1]
```

## Plots for Lag comparison

```{r}
color = rainbow(3)

for (year in 1:15) {
  par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
  
  Lag = "no_lag"
      
  plot(t(get(paste0("heatmap_data_", Lag, "_Spearman")))[,year],
       ylim = c(min(get(paste0("heatmap_data_", Lag, "_Spearman"))),
                max(get(paste0("heatmap_data_", Lag, "_Spearman")))),
       pch = 19,
       xlab = "provinces",
       ylab = paste(Correlation_method[i],"Correlation"),
       col = color[1],
       main = paste(Correlation_method[i], year+2005),
      xaxt = "n")
  
  axis(side = 1, at = 1:76)#, labels = province_names[-which(province_names[,1] == "Bungkan"), 1])
  
  for (Lag in c("1_month_lag", "2_month_lag")) {
    i = which(Lag == c("1_month_lag", "2_month_lag"))
    points(t(get(paste0("heatmap_data_", Lag, "_Spearman")))[,year],
         pch = 19,
         col = color[i+1],
         xaxt = "n")
    
    #axis(side = 1, at = 1:76, labels = province_names[-which(province_names[,1] == "Bungkan"), 1])    
  }
  
  legend("topright",
           legend = c("no Lag","1 month Lag", "2 month Lag"),
           pch = 19,
           col = color,
           inset=c(-0.3,0))
}
```

### Lag plots for every region

```{r}
color = rainbow(3)

for (region in regions) {
  for (year in 1:15) {
    par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
    
    Lag = "no_lag"
        
    plot(t(get(paste0("heatmap_data_", Lag, "_Spearman")))[which(colnames(get(paste0("heatmap_data_", Lag, "_Spearman"))) %in% get(paste0("provinces_", region))), year],
         
         ylim = c(min(get(paste0("heatmap_data_", Lag, "_Spearman"))),
                  max(get(paste0("heatmap_data_", Lag, "_Spearman")))),
         pch = 19,
         xlab = "provinces",
         ylab = paste(Correlation_method[i],"Correlation"),
         col = color[1],
         main = paste(Correlation_method[i], "correlation,", year+2005, "in the", region, "region"),
        xaxt = "n")
    
    axis(side = 1, at = 1:length(get(paste0("provinces_", region))), labels = get(paste0("provinces_", region)))
    
    for (Lag in c("1_month_lag", "2_month_lag")) {
      i = which(Lag == c("1_month_lag", "2_month_lag"))
      points(t(get(paste0("heatmap_data_", Lag, "_Spearman")))[which(colnames(get(paste0("heatmap_data_", Lag, "_Spearman"))) %in% get(paste0("provinces_", region))),year],
           pch = 19,
           col = color[i+1],
           xaxt = "n")
      
      #axis(side = 1, at = 1:76, labels = province_names[-which(province_names[,1] == "Bungkan"), 1])    
    }
    
    legend("topright",
             legend = c("no Lag","1 month Lag", "2 month Lag"),
             pch = 19,
             col = color,
             inset=c(-0.3,0))
  }  
}
```

# Total correlation over full time period (2006-2020)
```{r}
b = 5:182

for (p in 1:76) {
  
  for (i in 0:2) {
    
    Pearson = cor(x = as.vector(Precipitation[p,b], mode = "double"),
               y = as.vector(Incidence_combined[p,(b-2+i)], mode = "double"),
               method = "pearson")
  
    Spearman = cor(x = as.vector(Precipitation[p,b], mode = "double"),
                y = as.vector(Incidence_combined[p,(b-2+i)], mode = "double"),
                method = "spearman")
    
    if (i == 0) {
      Corr_complete = data.frame(c(Pearson, Spearman))
    }
    else {
      Corr_complete = cbind(Corr_complete, c(Pearson, Spearman))
    }
  }
  
  Corr_complete = cbind(paste0(Incidence_combined[p,2]), c("Pearson", "Spearman"), Corr_complete)
  
  colnames(Corr_complete) = c("Province", "Correlation method", "no Lag", "1 month Lag", "2 month Lag")
  
  if (p == 1) {
    correlation_total = Corr_complete 
  }
  else {
    correlation_total = rbind(correlation_total, Corr_complete)
  }
}

#add a region column
regions = c("central", "north", "northeastern", "south")
Region = character()

for (p in 1:152) {
  for (i in 1:4) {
    if (correlation_total[p,1] %in% get(paste0("provinces_", regions[i]))) {
      Region = c(Region, paste0(regions[i]))
    }
  }
}

correlation_total = cbind(correlation_total, Region)
```

### Data for tempo-spatial map

```{r}
correlation_map = correlation_total[which(correlation_total[,2] == "Spearman"), c(1,3)]
colnames(correlation_map) = c("Districts_names", "correlation")
```

### Correlation plots per region from 2006-2020
```{r}
for (Lag in 3:5) {
  Correlation_method = c("Pearson", "Spearman")
  
  for (i in 1:2) {
    par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
    
    plot(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag],
         ylim = c(min(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag]),
                  max(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag])),
         pch = 19,
         xlab = "Province (sorted by code)",
         ylab = paste(Correlation_method[i],"Correlation"),
         col = factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))]),
         main = paste(Correlation_method[i],
                      colnames(correlation_total[Lag])));
    
    legend("topright",
           legend = levels(factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))])),
           pch = 19,
           col = factor(levels(factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))]))),
           inset=c(-0.3,0))  
  }  
}


```

### Code to save plot as .png file
```{r}

Lag = 3
i = 2

png(file = "correlation_plot.png", width = 2400, height = 2400,res = 300)

plot(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag],
     ylim = c(min(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag]),
              max(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag])),
     pch = 19,
     xlab = "Province (sorted by official code)",
     ylab = paste(Correlation_method[i],"correlation coefficient"),
     col = factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))]),
     main = paste(Correlation_method[i], "correlation with no lag"));

legend("topright",
       legend = levels(factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))])),
       pch = 19,
       col = factor(levels(factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))]))))


dev.off()
```
```{r}

Lag = 4
i = 2

png(file = "correlation_plot_1_lag.png", width = 2400, height = 2400,res = 300)

plot(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag],
     ylim = c(min(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag]),
              max(correlation_total[which(correlation_total[,2] == paste(Correlation_method[i])),Lag])),
     pch = 19,
     xlab = "Province (sorted by official code)",
     ylab = paste(Correlation_method[i],"correlation coefficient"),
     col = factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))]),
     main = paste(Correlation_method[i], "correlation with 1-month lag"));

legend("topright",
       legend = levels(factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))])),
       pch = 19,
       col = factor(levels(factor(Region[which(correlation_total[,2] == paste(Correlation_method[i]))]))))


dev.off()
```

### Central regions provinces sorted by latitude

```{r}
correlation_total_spear = correlation_total[which(correlation_total[,2] == "Spearman"),]
correlation_total_spear[which(correlation_total_spear[,6] == "central"),]

Precipitation[which(Precipitation[,2] %in% provinces_central),c(2,4)]


plot(x = rank(correlation_total_spear[which(correlation_total_spear[,6] == "central"),3]), y = rank(Precipitation[which(Precipitation[,2] %in% provinces_central),4]), pch = 19)
```


## Comparing Lag for every region (Plots)
```{r}
for (reg in regions) {
  Correlation_method = c("Pearson", "Spearman")
  correlation_per_region = correlation_total[which(correlation_total[,6] == reg),]
  Lag = colnames(correlation_per_region[3:5]) 
  
  for (i in Correlation_method) {
    par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
    
    plot(correlation_per_region[which(correlation_per_region[,2] == paste(i)), 3],
         ylim = c(min(correlation_per_region[which(correlation_per_region[,2] == paste(i)),3:5]),
                  max(correlation_per_region[which(correlation_per_region[,2] == paste(i)),3:5])),
         pch = 19,
         xlab = "Province (sorted by code)",
         ylab = paste(i,"Correlation"),
         col = "black",
         main = paste(i,
                      correlation_per_region[1,6],
                      "region"))
    points(correlation_per_region[which(correlation_per_region[,2] == paste(i)), 4],
           pch = 19,
           col = "red")
    
    points(correlation_per_region[which(correlation_per_region[,2] == paste(i)), 5],
           pch = 19,
           col = "green")
    
    legend("topright",
           legend = Lag,
           pch = 19,
           col = c("black", "red", "green"),
           inset=c(-0.3,0))
  }
}
```

### save plot code

```{r}
i = "Spearman"

for (reg in regions) {
  
  region_names = c("central", "northern", "northeastern", "southern")
  
  Correlation_method = c("Pearson", "Spearman")
  correlation_per_region = correlation_total[which(correlation_total[,6] == reg),]
  Lag = colnames(correlation_per_region[3:5])
  
  png(file = paste0("Lag_comparison_", reg, ".png"), width = 2400, height = 2400, res = 300)
  
  plot(correlation_per_region[which(correlation_per_region[,2] == paste(i)), 3],
         ylim = c(min(correlation_per_region[which(correlation_per_region[,2] == paste(i)),3:5]),
                  max(correlation_per_region[which(correlation_per_region[,2] == paste(i)),3:5])),
         pch = 19,
         xlab = paste(region_names[which(regions %in% reg)], "provinces (sorted by code)"),
         ylab = paste(i,"Correlation"),
         col = "black",
         main = paste(i,
                      correlation_per_region[1,6],
                      "region"))
  points(correlation_per_region[which(correlation_per_region[,2] == paste(i)), 4],
           pch = 19,
           col = "red")
    
  points(correlation_per_region[which(correlation_per_region[,2] == paste(i)), 5],
         pch = 19,
         col = "green")
  
  if (reg == "central") {
    legend("topright",
           legend = Lag,
           pch = 19,
           col = c("black", "red", "green"))
  }  
  dev.off()
}
```


# Regional correlation

## Incidence table

```{r}
Incidence_regions = NULL

for (region in regions) {
  
  Dengue_region = data.frame(paste0(region))
  
  regional_Population = data.frame(t(apply(Population[which(Population[,2] %in% get(paste0("provinces_", region))), 3:17], 2, sum)))
  colnames(regional_Population) = 2006:2020
  
  for (year in 2006:2020) {
    b = year - 2005
    
    data = get(paste0("Dengue_", year))
    subset_data = data[which(data[,2] %in% get(paste0("provinces_", region))), 4:15]
    sum_data = data.frame(t(apply(subset_data, 2, sum)))
    
    sum_data = sum_data / regional_Population[1,b] * 100000
    
    Dengue_region = cbind(Dengue_region, sum_data)
  }
  
  Incidence_regions = rbind(Incidence_regions, Dengue_region)
}

colnames(Incidence_regions) = c("Region", colnames(Dengue_combined)[3:182])

```

## Precipitation table
```{r}
Precipitation_regions = NULL

for (region in regions) {
  
  Prec_region = data.frame(paste0(region))
  
  subdata = Precipitation[which(Precipitation[,2] %in% get(paste0("provinces_", region))), c(2, 5:184)]
  
  subarea = area[which(area[,2] %in% get(paste0("provinces_", region))), 3]
  
  litre_precipitation = apply(subdata[,2:181], 2, function(x){sum(x * subarea)})
  
  regional_area = sum(area[which(area[,2] %in% get(paste0("provinces_", region))), 3])
  
  Prec_region = cbind(Prec_region, data.frame(t(litre_precipitation / regional_area)))
  
  Precipitation_regions = rbind(Precipitation_regions, Prec_region)
}

colnames(Precipitation_regions) = c("Region", colnames(Dengue_combined)[3:182])
```

## Actual Correlation
```{r}
b = 2:179

for (p in 1:4) {
  
  for (i in 0:2) {
    
    Pearson = cor(x = as.vector(Precipitation_regions[p,b], mode = "double"),
               y = as.vector(Incidence_regions[p,(b+i)], mode = "double"),
               method = "pearson")
  
    Spearman = cor(x = as.vector(Precipitation_regions[p,b], mode = "double"),
                y = as.vector(Incidence_regions[p,(b+i)], mode = "double"),
                method = "spearman")
    
    if (i == 0) {
      Corr_complete = data.frame(c(Pearson, Spearman))
    }
    else {
      Corr_complete = cbind(Corr_complete, c(Pearson, Spearman))
    }
  }
  
  Corr_complete = cbind(paste0(Incidence_regions[p,1]), c("Pearson", "Spearman"), Corr_complete)
  
  colnames(Corr_complete) = c("region", "Correlation method", "no Lag", "1 month Lag", "2 month Lag")
  
  if (p == 1) {
    regional_correlation = Corr_complete 
  }
  else {
    regional_correlation = rbind(regional_correlation, Corr_complete)
  }
}
```

# Jun-Sep monsoon data

```{r}
a = 10:13
d = NULL

for (b in 0:14) {
  c = a+12*b
  d = c(d, c)
}

monsoon_prec = Precipitation[,c(1:2, d)]

monsoon_inc_0_lag = Incidence_combined[,c(1:2, d-2)]
monsoon_inc_1_lag = Incidence_combined[, c(1:2, d-1)]
monsoon_inc_2_lag = Incidence_combined[, c(1:2, d)]

```




