---
title: "Thailand Precipitation and Dengue Analysis"
output: pdf_document
date: "2023-05-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Upload

## Precipitation data
```{r}
library(readxl)
Precipitation <- as.data.frame(read_excel("ClimateData_2006_2020.xlsx", sheet = "era5_tp_2006_2020_thailand_monm"))
Precipitation
```

## Dengue data
```{r}
library(readxl)
Dengue_2006 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2006"))
Dengue_2007 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2007"))
Dengue_2008 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2008"))
Dengue_2009 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2009"))
Dengue_2010 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2010"))
Dengue_2011 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2011"))
Dengue_2012 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2012"))
Dengue_2013 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2013"))
Dengue_2014 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2014"))
Dengue_2015 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2015"))
Dengue_2016 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2016"))
Dengue_2017 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2017"))
Dengue_2018 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2018"))
Dengue_2019 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2019"))
Dengue_2020 <- as.data.frame(read_excel("dengue_data.xlsx", sheet = "2020"))

Dengue = list(Dengue_2006, Dengue_2007, Dengue_2008, Dengue_2009, Dengue_2010, Dengue_2011, Dengue_2012, Dengue_2013, Dengue_2014, Dengue_2015, Dengue_2016, Dengue_2017, Dengue_2018, Dengue_2019, Dengue_2020)
```

## Population Data
```{r}
Population <- as.data.frame(read_excel("population.xlsx", sheet = "Tabelle1"))
```


## Each province gets its code (number)
```{r}
library(readxl)
province_numbers <- as.data.frame(read_excel("province_numbers.xlsx"))
province_numbers = province_numbers[-27,]
```

# Data Copies for safety reasons 

## Precipitation
```{r}
Precipitationc = Precipitation
```

## Dengue Fever
```{r}
Dengue_2006c = Dengue_2006
Dengue_2007c = Dengue_2007
Dengue_2008c = Dengue_2008
Dengue_2009c = Dengue_2009
Dengue_2010c = Dengue_2010
Dengue_2011c = Dengue_2011
Dengue_2012c = Dengue_2012
Dengue_2013c = Dengue_2013
Dengue_2014c = Dengue_2014
Dengue_2015c = Dengue_2015
Dengue_2016c = Dengue_2016
Dengue_2017c = Dengue_2017
Dengue_2018c = Dengue_2018
Dengue_2019c = Dengue_2019
Dengue_2020c = Dengue_2020
```

## Province numbers
```{r}
province_numbersc = province_numbers
```

## Population 
```{r}
Populationc = Population
```

# Data Cleaning

## list of province names
```{r}
province_names = c("Bangkok", "Samut Prakan", "Nonthaburi", "Pathum Thani", "Phra Nakhon Si Ayutthaya", "Ang Thong", "Lop Buri", "Sing Buri", "Chai Nat", "Saraburi", "Chon Buri", "Rayong", "Chanthaburi", "Trat", "Chachoengsao", "Prachin Buri", "Nakhon Nayok", "Sa Kaeo", "Nakhon Ratchasima", "Buri Ram", "Surin", "Si Sa Ket", "Ubon Ratchathani", "Yasothon", "Chaiyaphum", "Amnat Charoen", "Bungkan", "Nong Bua Lam Phu", "Khon Kaen", "Udon Thani", "Loei", "Nong Khai", "Maha Sarakham", "Roi Et", "Kalasin", "Sakon Nakhon", "Nakhon Phanom", "Mukdahan", "Chiang Mai", "Lamphun",  "Lampang",  "Uttaradit", "Phrae", "Nan",    "Phayao", "Chiang Rai", "Mae Hong Son",  "Nakhon Sawan", "Uthai Thani", "Kamphaeng Phet", "Tak", "Sukhothai", "Phitsanulok", "Phichit", "Phetchabun", "Ratchaburi", "Kanchanaburi", "Suphan Buri", "Nakhon Pathom",  "Samut Sakhon", "Samut Songkhram", "Phetchaburi", "Prachuap Khiri Khan", "Nakhon Si Thammarat", "Krabi", "Phangnga", "Phuket", "Surat Thani", "Ranong", "Chumphon", "Songkhla", "Satun", "Trang", "Phatthalung", "Pattani", "Yala", "Narathiwat")

province_names = as.data.frame(province_names)
province_names
```

## Align province names, delete useless rows and add province numbers for all data sets

### Population Data

***Alignment of the province names***
```{r}
province_names[,1] %in% Population[,1] #is the province name in the Population table?
which((province_names[,1] %in% Population[,1]) == FALSE) #which provinces are NOT in the Population table?
province_names[which((province_names[,1] %in% Population[,1]) == FALSE),1] #what are the names of these "missing" provinces

Population[31,1] = province_names[5,1]
```

***Deletion of useless rows***
```{r}
Population = Population[which((Population[,1] %in% province_names[,1])), ]
rownames(Population) = c(1:77)
Population
```

***Addition of province numbers and sort rows by number***
```{r}
Population_1 = Population

for (x in 1:76)
  Population[x,] = Population_1[which(province_numbers[x,2] == Population_1[,1]),]

Population = cbind(province_numbers$Number, Population)
colnames(Population)[1] = "Code number"
```
### Precipitation Data

***Deletion of Number Column***
```{r}
Precipitation = Precipitation[,-1]
```

***Alignment of the province names***
```{r}
province_names[,1] %in% Precipitation[,1] #is the province name in the Precipitation table?
which((province_names[,1] %in% Precipitation[,1]) == FALSE) #which provinces are NOT in the Precipitation table?
province_names[which((province_names[,1] %in% Precipitation[,1]) == FALSE),1] #what are the names of these "missing" provinces
```

***Deletion of useless rows***
```{r}
Precipitation = Precipitation[which((Precipitation[,1] %in% province_names[,1])), ]
rownames(Precipitation) = c(1:77)
Precipitation
```

***Addition of province numbers and sort rows by number***
```{r}
#NICHT ausführen --> noch nicht fertig
Precipiation_1 = Precipitation

for (x in 1:76)
  Precipitation[x,] = Precipiation_1[which(province_numbers[x,2] == Precipiation_1[,1]),]

Precipitation = cbind(province_numbers$Number, Precipitation)
colnames(Precipitation)[1] = "Code number"
```

### Dengue 2006

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2006[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2006[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2006[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2006[28,1] = province_names[5,1]
```

***Deletion of useless rows***
```{r}
Dengue_2006 = Dengue_2006[which((Dengue_2006[,1] %in% province_names[,1])), ]
rownames(Dengue_2006) = c(1:76)
Dengue_2006
```

***Addition of province numbers and sort rows by number***
```{r}
Dengue_1 = Dengue_2006

for (x in 1:76)
  Dengue_2006[x,] = Dengue_1[which(province_numbers[x,2] == Dengue_1[,1]),]

Dengue_2006 = cbind(province_numbers$Number, Dengue_2006)
colnames(Dengue_2006)[1] = "Code number"
```

### Dengue 2007

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2007[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2007[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2007[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2007[28,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2007 = Dengue_2007[which((Dengue_2007[,1] %in% province_names[,1])), ]
rownames(Dengue_2007) = c(1:76)
Dengue_2007
```

***Addition of province numbers and sort rows by number***
```{r}
Dengue_1 = Dengue_2007

for (x in 1:76)
  Dengue_2007[x,] = Dengue_1[which(province_numbers[x,2] == Dengue_1[,1]),]

Dengue_2007 = cbind(province_numbers$Number, Dengue_2007)
colnames(Dengue_2007)[1] = "Code number"
```

### Dengue 2008

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2008[,1]
which((province_names[,1] %in% Dengue_2008[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2008[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2008[6,1] = province_names[5,1] #set the correct name for the province
Dengue_2008[64,1] = province_names[35,1]
```

***Deletion of useless rows***
```{r}
Dengue_2008 = Dengue_2008[which((Dengue_2008[,1] %in% province_names[,1])), ]
rownames(Dengue_2008) = c(1:76)
Dengue_2008
```

***Addition of province numbers and sort rows by number***
```{r}
Dengue_1 = Dengue_2008

for (x in 1:76)
  Dengue_2008[x,] = Dengue_1[which(province_numbers[x,2] == Dengue_1[,1]),]

Dengue_2008 = cbind(province_numbers$Number, Dengue_2008)
colnames(Dengue_2008)[1] = "Code number"
```

### Dengue 2009

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2009[,1]
which((province_names[,1] %in% Dengue_2009[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2009[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2009[6,1] = province_names[5,1] #set the correct name for the province
Dengue_2009[64,1] = province_names[35,1]
```

***Deletion of useless rows***
```{r}
Dengue_2009 = Dengue_2009[which((Dengue_2009[,1] %in% province_names[,1])), ]
rownames(Dengue_2009) = c(1:76)
Dengue_2009
```

***Addition of province numbers and sort rows by number***
```{r}
Dengue_1 = Dengue_2009

for (x in 1:76)
  Dengue_2009[x,] = Dengue_1[which(province_numbers[x,2] == Dengue_1[,1]),]

Dengue_2009 = cbind(province_numbers$Number, Dengue_2009)
colnames(Dengue_2009)[1] = "Code number"
```

### Dengue 2010

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2010[,1]
which((province_names[,1] %in% Dengue_2010[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2010[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2010[6,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2010 = Dengue_2010[which((Dengue_2010[,1] %in% province_names[,1])), ]
rownames(Dengue_2010) = c(1:76)
Dengue_2010
```

***Addition of province numbers and sort rows by number***
```{r}
Dengue_1 = Dengue_2010

for (x in 1:76)
  Dengue_2010[x,] = Dengue_1[which(province_numbers[x,2] == Dengue_1[,1]),]

Dengue_2010 = cbind(province_numbers$Number, Dengue_2010)
colnames(Dengue_2010)[1] = "Code number"
```

### Dengue 2011

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2011[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2011[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2011[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2011[6,1] = province_names[5,1] #set the correct name for the province
Dengue_2011[65,1] = province_names[35,1]
```

***Deletion of useless rows***
```{r}
Dengue_2011 = Dengue_2011[which((Dengue_2011[,1] %in% province_names[,1])), ]
rownames(Dengue_2011) = c(1:77)
Dengue_2011
```

### Dengue 2012

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2012[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2012[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2012[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2012[37,1] = province_names[5,1] #set the correct name for the province
Dengue_2012[14,1] = province_names[35,1]
```

***Deletion of useless rows***
```{r}
Dengue_2012 = Dengue_2012[which((Dengue_2012[,1] %in% province_names[,1])), ]
rownames(Dengue_2012) = c(1:77)
Dengue_2012
```

### Dengue 2013

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2013[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2013[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2013[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2013[37,1] = province_names[5,1]  #set the correct name for the province
Dengue_2013[14,1] = province_names[35,1]
```

***Deletion of useless rows***
```{r}
Dengue_2013 = Dengue_2013[which((Dengue_2013[,1] %in% province_names[,1])), ]
rownames(Dengue_2013) = c(1:77)
Dengue_2013
```

### Dengue 2014

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2014[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2014[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2014[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2014[37,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2014 = Dengue_2014[which((Dengue_2014[,1] %in% province_names[,1])), ]
rownames(Dengue_2014) = c(1:77)
Dengue_2014
```

### Dengue 2015

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2015[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2015[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2015[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2015[37,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2015 = Dengue_2015[which((Dengue_2015[,1] %in% province_names[,1])), ]
rownames(Dengue_2015) = c(1:77)
Dengue_2015
```

### Dengue 2016

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2016[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2016[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2016[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2016[37,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2016 = Dengue_2016[which((Dengue_2016[,1] %in% province_names[,1])), ]
rownames(Dengue_2016) = c(1:77)
Dengue_2016
```

### Dengue 2017

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2017[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2017[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2017[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2017[31,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2017 = Dengue_2017[which((Dengue_2017[,1] %in% province_names[,1])), ]
rownames(Dengue_2017) = c(1:77)
Dengue_2017
```

### Dengue 2018

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2018[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2018[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2018[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2018[31,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2018 = Dengue_2018[which((Dengue_2018[,1] %in% province_names[,1])), ]
rownames(Dengue_2018) = c(1:77)
Dengue_2018
```

### Dengue 2019

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2019[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2019[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2019[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2019[31,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2019 = Dengue_2019[which((Dengue_2019[,1] %in% province_names[,1])), ]
rownames(Dengue_2019) = c(1:77)
Dengue_2019
```

### Dengue 2020

***Alignment of the province names***
```{r}
province_names[,1] %in% Dengue_2020[,1] #is the province name in the Dengue table?
which((province_names[,1] %in% Dengue_2020[,1]) == FALSE) #which provinces are NOT in the Dengue table?
province_names[which((province_names[,1] %in% Dengue_2020[,1]) == FALSE),1] #what are the names of these "missing" provinces

Dengue_2020[31,1] = province_names[5,1] #set the correct name for the province
```

***Deletion of useless rows***
```{r}
Dengue_2020 = Dengue_2020[which((Dengue_2020[,1] %in% province_names[,1])), ]
rownames(Dengue_2020) = c(1:77)
Dengue_2020
```


